Switch equivalence and weak prime domains for fusions

### 2024-11-22 full_version/leftlinear.tex
minori

\chapter{Left-linear DPO-rewriting systems}

$\mathcal{M}$-adhesive categories are the right context in which to perform abstract rewriting using the so-called ``douple pushout approach'' (DPO). We will recall the basic definitions and properties of this approach to abstract rewriting. 
\section{First definitions and properties}
We are now going to study rewriting systems in $\mathcal{M}$-adhesive categories.

\begin{definition}[\cite{habel2012mathcal,heindel2009category}]
	Let $\X$ be an $\mathcal{M}$-adhesive category, a  \emph{left $\mathcal{M}$-linear} rule $\rho$ is a pair $(l,r)$ of arrows with the same domain, such that $l$ belongs to $\mathcal{M}$.  The rule $\rho$ is said to be \emph{$\mathcal{M}$-linear} if $r\in \mathcal{M}$ too. A rule $\rho$ is said to be \emph{consuming} if $l$ is not an isomorphism. We will represent a rule $\rho$ as a span 
	\[\xymatrix{L & K\ar[l]_{l} \ar[r]^{r} & R}\]
$L$ is the \emph{left-hand side}, $R$ is the \emph{right-hand side} and $K$ the \emph{glueing object}.

A \emph{left-linear DPO-rewriting system} is a pair $(\X, \R)$ where $\X$ is a $\mathcal{M}$-adhesive category and $R$ is a set of left $\mathcal{M}$-linear rules. $(\X, \R)$ will be called \emph{linear} if every rule in $R$ is $\mathcal{M}$-linear. Similarly, a \emph{consuming} left-linear DPO-rewriting system is one in which every rule is consuming.

Given  two objects $G$ and $H$ and a rule $\rho=(l,r)$ in $\R$, a \emph{direct derivation $\mathscr{D}$ from $G$ to $H$ applying the rule $\rho$}, is a diagram as the one below, in which both squares are pushouts. 
	\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{h}\\G & \ar[l]^{f} D \ar[r]_{g}& H}\]
	The arrow $n$ is called the \emph{match} of the derivation, while $h$ is its \emph{back-match}.
	We will denote a direct derivation $\dder{D}$ between $G$ and $H$ as $\dder{D}\colon G\Mapsto H$. 
\end{definition}

\begin{example}\todo{
esempi di	derivazione}
\end{example}

\begin{remark}\label{exa:conc} Let  $\dder{D}\colon G\Mapsto H$ be the direct derivation 
		\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{h}\\G & \ar[l]^{f} D \ar[r]_{g}& H}\]
	If $\phi\colon G'\to G$ and $\psi\colon H\to H'$ are two isomorphisms, 	we can consider the direct derivation	$\phi * \dder{D}*\psi \colon G'\Mapsto H'$ given by the following diagram.
	\[\xymatrix{L \ar[d]_{\phi^{-1} \circ n}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{\psi \circ h}\\G' & \ar[l]^{\phi^{-1} \circ f} D \ar[r]_{\psi \circ g}& H'}\]

	In particular, we will use $\phi*\dder{D}$ and $\dder{D}*\psi$  to denote $\phi*\dder{D}*\id{H}$ and $\id{G}*\dder{D}*\psi$.
\end{remark}

$\mathcal{M}$-adhesivity of $\X$ guarantes the essential uniqueness of the result obtained rewriting an object, as shown by the next proposition.

\begin{proposition}\label{prop:unique} Let $\X$  be a $\mathcal{M}$-adhesive category. Suppose that the two direct derivations $\mathscr{D}$ and $\mathscr{D'}$ below, with the same match and applying the same left $\mathcal{M}$-linear rule $\rho$ are given.
	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} & L \ar[d]_{m}& K \ar[d]^{k'}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h'}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H & G & \ar@{>->}[l]^{f'} D' \ar[r]_{g'}& H'}\]
Then there exist isomorphisms $t\colon D\to D'$ and $s\colon H\to H'$ as in the following diagram.
\[\xymatrix@C=40pt{&&D' \ar[r]^{g'} \ar@{>->}@/_.45cm/[dll]_{f'}&H'\\G & L \ar[l]_{m} & K \ar[u]_{k'} \ar[d]^{k}\ar[r]^{r} \ar@{>->}[l]_{l} &R\ar[u]^{h'} \ar[d]_{h}\\&&D\ar@{>->}@/^.45cm/[ull]^{f}\ar@/^.4cm/@{.>}[uu]^(.4){t}|\hole \ar[r]_{g}&H\ar@/_.4cm/@{.>}[uu]_{s}}\]
\end{proposition}
\begin{proof}
	By construction, the pairs $(k, f)$ and $(k', f')$ are pushout complements of $l$ and $n$. Thus, the existence of the isomorphism $t\colon D\to D'$ follows from \Cref{lem:pocomp}. Now, computing we have
	\begin{align*}
		g'\circ t \circ k &= g' \circ k'\\&=h'\circ r
	\end{align*}
	Hence, we have the wanted $s\colon H\to H'$. To see that $s$ is an isomorphism, consider the diagram 
	\[\xymatrix{K  \ar@/^.4cm/[rr]^{k'}\ar[d]_{r} \ar[r]_{k}& \ar[r]_{t} D \ar[d]^{g}& D' \ar[d]^{g'}\\ R \ar@/_.4cm/[rr]_{h'} \ar[r]^{h}& H \ar[r]^{s}& H'}\]
	By hypothesis the whole rectangle and its left half are pushouts, therefore, by \Cref{lem:po1} its right square is a pushout too. The claim now follows from the fact that the pushout of an isomorphism is an isomorphism.
\end{proof}

If we look to direct derivations as transitions, it is natural to consider them as edges in a direct graph. Taking objects as vertices objects led us to the following definition \cite{heindel2009category}.

\begin{definition}
	Let $(\X, \R)$ be a left-linear DPO-rewrityng system, with $\X$ $\mathcal{M}$-adhesive. The \emph{DPO-derivation graph} of $(\X, \R)$ is the (large)  directed graph $\gpo$ having as vertices the objects of $\X$ and in which an edge between $G$ and $H$ is a direct derivation $\dder{D}\colon G\Mapsto H$.	A \emph{derivation} $\der{D}$ between two objects $G$ and $H$ is a path between them in $\gpo$. The \emph{source} and \emph{target} of $\der{D}$ are, respectively, $G$ and $H$.
\end{definition}

\begin{remark}
We can spell out more explicitly what  a derivation $\dder{D}$ is.  An \emph{empty derivation} starting and ending in $G$ is just $G$ itself.  A \emph{non-empty derivation} $\dder{D}$ is a sequence $\{\dder{D}_i\}_{i=0}^n$ of direct derivations such that:
\begin{enumerate}
	\item for every index $i$, $\dder{D}_i$ is a direct derivation $G_i \Mapsto G_{i+1}$;
	\item $G_0=G$ and $G_{n+1}=H$.
\end{enumerate}

We will call the number $n+1$ the \emph{length} of the derivation, denoted by $\lgh(\der{D})$. We will also say that an empty derivation has length $0$.

Moreover,  if every $\dder{D}_i$ applies the rule $\rho_i\in R$, then we can define an associated sequence of rules as $r(\der{D})$ as $\{\rho_i\}_{i=0}^n$.
\end{remark}

\begin{remark}\label{rem:func}
	Consider a derivation $\der{D}$ in a left-linear DPO-rewriting system $(\X, \R)$. We can take the subcategory $\Delta(\der{D})$ of $\X$ given by the arrows appearing in $\der{D}$. This subcategory comes equipped with an inclusion functor $I(\der{D})\colon \Delta(\der{D})\to \X$. Moreover, we can further define $\Deltamin(\der{D})$ as the subcategory of $\Delta(\der{D})$ containing only the bottom row of the derivation.
\end{remark}

\begin{notation}Let $\der{D}=\{\dder{D}_i\}_{i=0}^n$ be a derivation. We will depict the $i^\text{th}$ element $\dder{D}_i$ of $\der{D}$ as in the following diagram.  
	\[\xymatrix{L_i \ar[d]_{m_i}& K_i \ar[d]^{k_i}\ar[l]_{l_i} \ar[r]^{r_i} & R_i \ar[d]^{h_i} \\G_{i} & \ar[l]^{f_{i}} D_{i} \ar[r]_{g_{i}}& G_{i+1} }\]
	Notice that, in particular, if $\der{D}\colon G\to H$, then $G_0=G$ and $G_{n+1}=H$. When $\der{D}$ has length $1$ we will suppress the indexes. In such case, we will also identify $\der{D}$ with its only element. 
\end{notation}

\begin{example}\todo{esempi di derivazione}
\end{example}

\begin{definition}
	The \emph{DPO-derivation category} $\dpo$ of a left-linear DPO-rewriting system $(\X, \R)$ is the category in which arrows between $G$ and $H$ are given by, possibly empty, derivations. Composition is concatenation of paths in $\gpo$ and identities are given by empty derivations.
\end{definition} 	
\begin{remark}
	More explicitly, given $\der{D}=\{\dder{D}\}_{i=0}^n$ between $G$ and $H$ and $\der{D}'=\{\dder{D}'_i\}_{i=0}^m$, their concatenation $\der{D}\cdot\der{D}'$ is the derivation $\{\dder{E}_i\}_{i=0}^{m+n+1}$ in which
	\[\dder{E}_i:=\begin{cases}
		\dder{D}_i & i \leq n\\
		\dder{D}'_{i-(n+1)} & n< i 
	\end{cases}\]

Notice, moreover that, $\der{D}\cdot \der{D'}$ is equal to $\der{D}'$ if $\der{D}$ is empty, while it coincides with $\der{D}$ if $\der{D}'$ has length zero.
\end{remark}

 \Cref{exa:conc} allows us to compose derivations with isomorphisms.

\begin{definition} Let $(\X, \R)$ be a a left-linear DPO-rewriting system. Given a derivation $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ between $G$ and $H$ and isomorphisms $\phi\colon G\to G'$, $\psi\colon H\to H'$, the derivations  $\phi*\der{D}$ and $\der{D}*\psi$ are defined as
	\[\phi *\der{D} := \begin{cases}
		G' & \lgh(\der{D})=0\\ 
		\{\phi* \dder{D}_0\}\cdot \{\dder{D}_i\}_{i=1}^{n}  & \lgh(\der{D})\neq 0
	\end{cases} \qquad \der{D}*\psi := \begin{cases}
	H' & \lgh(\der{D})=0\\ 
	\{\dder{D}_i\}_{i=0}^{n-1} \cdot \{\dder{D}_n*\psi\} & \lgh(\der{D})\neq 0
	\end{cases}\]

	Moreover, if $\lgh(\der{D})>0$,  we define the derivation $\phi *\der{D} * \psi$ as
	\[\phi *\der{D} * \psi = \{\phi* \dder{D}_0\}\cdot \{\dder{D}_i\}_{i=1}^{n-1} \cdot \{\dder{D}_n*\psi\}\] 
\end{definition}

\begin{remark}
When $\der{D}$ consists only in the direct derivation $\dder{D}$, then $\phi*\der{D}*\psi$ is the derivation of length one whose unique element is $\phi*\dder{D}*\psi$.
\end{remark}

\subsection{Decorations and abstraction equivalence}
We are often interested in an object of $\X$ only up to isomorphism. It is therefore useful to consider a version of $\gpo$ in which vertices are classes of isomorphism of object of $\X$. In order to do so, some preliminary work is needed.

\begin{definition}\cite{mac2013categories}
Let $\X$ be a category, we say that  $\X$ is \emph{skeletal} if, for every two objects $X$ and $Y$, the existence of an isomorphism $\phi\colon X\to Y$ entails $X=Y$. A \emph{skeleton} for a category $\X$ is a full subcategory $\ske$ which is skeletal and such that the inclusion functor $\ske\to \X$ is an equivalence. 
\end{definition}

\begin{remark}
By definition the inclusion $\ske \to \X$ is an equivalence. In particular, this mean that, for every objects $X$ of $\X$ there exists $\pi(X)$ in $\ske$ and an isomorphism $\phi_X\colon \pi(X) \to X$.
\end{remark}

\begin{proposition}\label{prop:ske}
	Every category $\X$ has a skeleton. 
\end{proposition}
\begin{proof}
	For every object $X\in \X$, pick a single representative $\pi(X)$ of its isomorphism class. Let $\ske$ be the full subcategory given by these objects. By definition $\ske$ is skeletal and the inclusion functor is full, faithful and essentially surjective.\qedhere 
\end{proof}
\begin{remark}
	The proof of \Cref{prop:ske} relies on the axiom of choice for classes.
	\end{remark}
\begin{remark}
	It is possible to proof that every two skeleta of a given category $\X$ are isomorphic (not only equivalent). For the remaining of this paper we assume that a skeleton $\ske$ of $\X$ and a functor $\pi\colon \X\to \ske$ are chosen once and for all.
\end{remark}

\begin{definition}
	Let $(\X, \R)$ be a left-linear DPO-rewriting system, a \emph{decorated derivation} between two objects $G$ and $H$ is a triple $(\der{D}, \alpha, \omega)$, where $\der{D}$ is a derivation between $G$ and $H$, and $\alpha\colon \pi(G)\to G$ and $\omega\colon \pi(H)\to H$ are isomorphisms.
\end{definition}

\begin{notation}
	We will extend the use of the words length, source and target to decorated derivations in the obvious way, forgetting the decorations $\alpha$ and $\omega$.
\end{notation}

\begin{example}A decorated derivation $(\der{D}, \alpha, \omega)$ with $\der{D}$ empty is just a span
	\[\xymatrix{G & \pi(G) \ar[r]^-{\omega} \ar[l]_-{\alpha} & G}\]
	in which both $\omega$ and $\alpha$ are isomorphisms.
\end{example}

As we are interested in objects only up to isomorphism, so we are interested in (decorated) derivations only up to some notion of coherent isomorphism between them. This is done with the help of \Cref{rem:func}.

\begin{definition}Let $(\X, \R)$ be a left-linear DPO-rewriting system,  an \emph{abstraction equivalence} between two derivations $\der{D}$ and $\der{D'}$ with the same length and such that $r(\der{D})=r(\der{D}')$, is a family of isomorphisms $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ such that, for every $i\in [0, \lgh(\der{D})]$ the following diagram commutes
	\[\xymatrix@C=40pt{G'_i&D'_i \ar[r]^{g'_i} \ar[l]_{f'_i}&G'_{i+1}\\  L_i \ar[u]^{n'_i} \ar[d]_{n_i}& K_i \ar[u]^{k'_i} \ar[d]_{k_i} \ar[r]^{r_i} \ar[l]_{l_i} &R\ar[u]^{h'_i} \ar[d]_{h_i}\\G_i \ar@/_.45cm/[uu]_(.35){\phi_{G_i}}|\hole&D_i\ar[l]^{f_i}\ar@/_.45cm/[uu]_(.35){\phi_{D_i}}|\hole \ar[r]_{g_i}&G_{i+1}\ar@/_.45cm/[uu]_{\phi_{H_i}}}\]

Given two decorated derivations $(\der{D}, \alpha, \omega)$ and $(\der{D}', \alpha', \omega')$, we say that they are \emph{abstraction equivalent}, if $\lgh(\der{D})=\lgh(\der{D}')$, $r(\der{D})=r(\der{D'})$, and there exists an of abstraction equivalence between $\der{D}$ and  $\der{D}$ such that the triangles below commute.
\[\xymatrix@C=15pt{&\pi(G_0) \ar[dr]^{\alpha'} \ar[dl]_{\alpha}&&& \pi(G_{n+1}) \ar[dr]^{\omega'} \ar[dl]_{\omega}\\ G_0 \ar[rr]_{\phi_{G_0}} && G'_0 &G_{n+1} \ar[rr]_{\phi_{G_{n+1}}} && G'_{n+1} } \]
We will use $\equiv^a$ to denote the resulting relation.
\end{definition}

\begin{remark}\label{rem:equi}
	It is immediate to see that $\equiv^a$ is an equivalence relation. Indeed $(\der{D}, \alpha, \omega)$ is abstract equivalent to itself via the abstract equivalence with the identities as components. Furthermore, if  $\{\phi_X\}_{X\in \Deltamin(\der{D})}$  witnesses $(\der{D}, \alpha, \omega)\equiv_a (\der{D}', \alpha', \omega')$, then considering $\{\phi^{-1}_X\}_{X\in \Deltamin(\der{D})}$ shows $(\der{D}', \alpha', \omega)\equiv_a (\der{D}, \alpha, \omega)$. Finally, transitivity is assured composing abstract equivalences.

	We will denote by $[\der{D}, \alpha, \omega]_a$ is just the equivalence class of  $(\der{D}, \alpha, \omega)$.
	 Such equivalence classes will be called  \emph{abstract decorated derivation}.  
\end{remark}

>>>>

\chapter{Independence in rewriting}

In this chapter we discuss the notion of independence between two
consecutive rewriting steps, a fundamental ingredient of the rewriting
theory, which comes into play when viewing a sequence of steps as a
concurrent computation. We observe that moving from linear to
left-linear rules leads to the failure of various basic properties of
the classical notion of independence used in the linear setting and we
single out a framework in which these can be re-established, possibly
in a weakened form.

\section{Sequentially independent and switchable derivations }\label{subsec:switch}

Sequential independence is the canonical notion of independence in the DPO approach.

\begin{definition}[Sequential independence]
	\label{de:sequential-independence}
	Let $\X$ be an $\mathcal{M}$-adhesive category 
	and $(\X, \R)$ a left-linear rewriting system. 
	Let also
	$\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation of length $2$, with
	$\dder{D}_i$ using rule $\rho_i = (l_i,r_i)$.  We say that $\dder{D}_0$ and
	$\dder{D}_1$ are
	\emph{sequentially independent}
	if there is a pair of arrows
	$i_0\colon R_0\to D_1$ and $i_1\colon L_1\to D_0$ such that the
	following diagram commutes. In this case $(i_0,i_1)$ is called an  \emph{independence
		pair}.
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{.>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{.>}@/_.25cm/[dlll]^(.4){i_1}
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}&& G_1 && \ar@{>->}[ll]^{f_1} D_1
		\ar[rr]_{g_1}&& G_2}
	\]
\end{definition}

\begin{remark}\label{rem:seqabs}
Let $\{\phi_X\}_{X\in \Deltamin(\der{D})}$ be an abstraction equivalence between $\der{D}=\{\dder{D}_i\}_{i=0}^1$ and $\der{D}'=\{\dder{D}'_i\}_{i=0}^1$. If there is an independence pair between $(i_0, i_1)$ between $\dder{D}_0$ and $\dder{D}_1$, then $(\phi_{D_0}\circ i_1, \phi_{D_1}\circ i_0)$ is an independence pair between $\dder{D}'_0$ and $\dder{D}'_1$. Indeed we have
\[\begin{split}
g'_0\circ \phi_{D_0}\circ i_1&=\phi_{G_1}\circ g_0\circ i_1\\&=\phi_{G_1}\circ m_1\\&= m'_1
\end{split}\qquad \begin{split}
f'_1\circ \phi_{D_1}\circ i_0&=\phi_{G_1}\circ f_1\circ i_0\\&=\phi_{G_1}\circ h_0\\&= h'_1
\end{split}\]
\end{remark}

\begin{remark}\label{rem:uni}
	It is worth mentioning that if $(\X, \R)$ is a linear rewriting system then two derivations $\dder{D}_0$ and $\dder{D}_1$ can have at most one independence pair. Indeed if $(i_0,i_1)$ and $(i'_0, i'_1)$ are independence pairs, then
	\[g_0\circ i_1 = g_0\circ i'_1 \qquad f_1\circ i_0=f_1\circ i'_0\]
	But in linear systems $g_0$ and $f_1$ are both monos, entailing $i_0=i'_0$ and $i_1=i'_1$.
\end{remark}

Intuitively, the existence of the independence pair
captures the possibility of switching the application of the two rules:
$f_0 \circ i_1$ is a match for $\rho_1$ in $G_0$ and the
existence of $i_0 : R_0 \to D_1$ means that $\rho_1$ does not delete
items in the image of $K_0$, thus $\rho_0$ can be applied after
$\rho_1$. The following definition formalise what it means for a derivation to be the switch of another.

\begin{definition}[Switch] \label{def:switch}
	Let $\X$ be an $\mathcal{M}$-adhesive category 
	and $(\X, \R)$ a left-linear rewriting system. 
	Let also
	$\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation made
	of two sequentially independent derivations $\dder{D}_0$,
	$\dder{D}_1$ with independence pair $(i_0, i_1)$, as in the diagram
	on the left below. A \emph{switch} of $\der{D}$ along $(i_0,i_1)$ is
	a derivation $\der{E}=\{\dder{E}_i\}_{i=0}^1$, between the same objects and using the same rules
	in reverse order, as in the diagram on the right below
	\[
	\def\objectstyle{\scriptstyle}\def\labelstyle{\scriptstyle}
	\xymatrix@C=1.1mm@R=8mm{
		{L_0} \ar[d]_{m_0}
		& & {K_0} \ar@{>->}[ll]_{l_0} \ar[rr]^{r_0} \ar[d]_{k_0}
		& & {R_0} \ar[drr]|(.35)\hole_(.6){h_0}  \ar@/^.22cm/[drrrrrr]|(.3)\hole^(.75){i_0}
		& & & & 
		% 
		{L_1}\ar[dll]|(.35)\hole^(.6){m_1} \ar@/_.22cm/_(.75){i_1}[dllllll]
		& & {K_1} \ar@{>->}[ll]_{l_1} \ar[rr]^{r_1} \ar[d]^{k_1}
		& & {R_1} \ar[d]^{h_1} \\
		% 
		% 
		{G_0}
		& & {D_0} \ar@{>->}[ll]^{f_0} \ar[rrrr]_{g_0}
		& & & & {G_1} & &
		& &  {D_1} \ar@{>->}[llll]^{f_1} \ar[rr]_{g_1}
		& & {G_2}
	}
	% 
	\xymatrix@C=1.1mm@R=8mm{
		{L_1} \ar[d]_{m_0'}
		& & {K_1} \ar@{>->}[ll]_{l_1} \ar[rr]^{r_1} \ar[d]_{k_0'}
		& & {R_1} \ar[drr]|(.35)\hole_(.6){h_0'}  \ar@/^.22cm/@{.>}^(.75){i_0'}|(.3)\hole[drrrrrr]
		& & & & 
		% 
		{L_0}\ar[dll]|(.35)\hole^(.6){m_1'} \ar@/_.22cm/@{.>}_(.75){i_1'}[dllllll] 
		& & {K_0} \ar@{>->}[ll]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\
		% 
		% 
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rrrr]_{g_0'}
		& & & & {G'_1} & &
		& &  {D_1'} \ar@{>->}[llll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	%
	\]
	such that there is an independence pair $(i_0', i_1')$ between
	$\dder{E}_0$ and $\dder{E}_1$ and 
	\begin{center}   
		$m_0=f_0' \circ i_1'$
		\qquad $h_1=g_1' \circ i_0'$
		\qquad $m_0'= f_0 \circ i_1$
		\qquad $h_1'= g_{1}\circ i_0$.
	\end{center}

	If a switch of $\dder{D}_0$ and $\dder{D}_1$ exists we say that
	they are \emph{switchable}.
\end{definition}

\begin{example}	\label{ex:seq-ind}
	Consider the rewriting system  $(\cat{Graph}, \R)$ and the derivarion $\der{D}$ introduced in \Cref{ex:1}.	Steps $\dder{D}_1$ and $\dder{D}_2$ are clearly sequential
	independent via the arrows $R_1 \to D_2$, $L_2 \to D_1$
	mapping nodes according to their numbering.
	%
	A switch of $\der{D}$ along such independence pair is the derivation
	$\der{E}$ depicted below. Instead,  the first two
	steps $\dder{D}_0$ and $\dder{D}_1$ are not sequential independent. Intuitively, this happens because $\rho_1$ uses the node generated by $\rho_0$ to produce a self-loop.

\begin{center}
	\begin{tikzpicture}[node distance=2mm, font=\small,  baseline=(current bounding box.center)]      
		\node (L1) at (0,2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {};
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of L1] (K1) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {};
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		%     \node [above=of K1] {$\rho_0$};
		\node [above] at (K1.north) {$\rho_0$};
		\node [right=of K1](R1) {
			\begin{tikzpicture}
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {};
				\node at (.5,0) [node, label=below:$2$] (2) {}; 
				\draw[coloredge] (1) to[out=20, in=160] (2);
				% 
				\pgfBox
			\end{tikzpicture}
		};
		\path (K1) edge[->] node[trans, above] {} (L1);
		\path (K1) edge[->] node[trans, above] {} (R1);

		\node at (4,2) (L3) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				\node at (0.5,0) [node, label=below:$2$] (2) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of L3] (K3) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				\node at (0.5,0) [node, label=below:$2$] (2) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		%     \node [above=of K3] {$\rho_2$};
		\node [above] at (K3.north) {$\rho_2$};
		\node [right=of K3] (R3) {
			\begin{tikzpicture}
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$12$] (12) {};
				% 
				\pgfBox
			\end{tikzpicture}
		};
		\path (K3) edge[->] node[trans, above] {} (L3);
		\path (K3) edge[->] node[trans, above] {} (R3);

		\node at (8,2) (L2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$2$] (1) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of L2] (K2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$2$] (1) {};
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		%     \node [above=of K2] {$\rho_1$};
		\node [above] at (K2.north) {$\rho_1$};

		\node [right=of K2] (R2) {
			\begin{tikzpicture}
				\node at (0,0) [node, label=below:$2$] (1) {}
				edge [in=55, out=85, loop] ();
				% 
				\pgfBox
			\end{tikzpicture}
		};
		\path (K2) edge[->] node[trans, above] {} (L2);
		\path (K2) edge[->] node[trans, above] {} (R2);

		%%%%%% second row
		\node at (0,0) (G1) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {};
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of G1] (D1) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {};
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node at (3,0) (G2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				\node at (0.5,0) [node, label=below:$2$] (2) {} ;
				\draw[coloredge] (1) to[out=20, in=160] (2);
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\path (D1) edge[->] node[trans, above] {} (G1);
		\path (D1) edge[->] node[trans, above] {} (G2);
		\path (L1) edge[->] node[trans, above] {} (G1);
		\path (K1) edge[->] node[trans, above] {} (D1);
		\path (R1) edge[->] node[trans, above] {} (G2);

		\node at (5.5,0) (D2) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {};
				\node at (0,0) [node, label=below:$1$] (1) {} ;
				\node at (0.5,0) [node, label=below:$2$] (2) {} ;
				\draw[coloredge] (1) to[out=20, in=160] (2);
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node at (7.4,0) (G3) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$12$] (12) {}
				edge [in=125, out=155, colorloop] ();
				% 
				\pgfBox
			\end{tikzpicture} 
		};

		\path (D2) edge[->] node[trans, above] {} (G2);
		\path (D2) edge[->] node[trans, above] {} (G3);
		\path (L3) edge[->] node[trans, above] {} (G2);
		\path (K3) edge[->] node[trans, above] {} (D2);
		\path (R3) edge[->] node[trans, above] {} (G3);

		\node at (9.1,0) (D3) {
			\begin{tikzpicture}
				% 
				\node at (0,0.53) {}; 
				\node at (0,0) [node, label=below:$12$] (12) {}
				edge [in=125, out=155, colorloop] ();
				% 
				\pgfBox
			\end{tikzpicture} 
		};
		\node [right=of D3] (G4) {
			\begin{tikzpicture}
				% 
				\node at (0,0) [node, label=below:$12$] (12) {}
				edge [in=55, out=85, loop] ()
				edge [in=125, out=155, colorloop] ();
				% 
				\pgfBox
			\end{tikzpicture} 
		};

		\node[font=\scriptsize, below] at (G1.south) {$G_0$};
		\node[font=\scriptsize, below] at (D1.south) {$D_0$};      
		\node[font=\scriptsize, below] at (G2.south) {$G_1$};
		\node[font=\scriptsize, below] at (D2.south) {$D_1'$};      
		\node[font=\scriptsize, below] at (G3.south) {$G_2'$};
		\node[font=\scriptsize, below] at (D3.south) {$D_2'$};      
		\node[font=\scriptsize, below] at (G4.south) {$G_3$};

		\path (D3) edge[->] node[trans, above] {} (G3);
		\path (D3) edge[->] node[trans, above] {} (G4);
		\path (L2) edge[->] node[trans, above] {} (G3);
		\path (K2) edge[->] node[trans, above] {} (D3);
		\path (R2) edge[->] node[trans, above] {} (G4);
	\end{tikzpicture}
	%
\end{center}

\end{example}

We can now prove some properties relating switching and abstraction equivalence. As a first step we can show that switches along the same independence paire are unique up to abstraction equivalence.

\begin{lemma}[Uniqueness of switches]
	\label{thm:switch_uni}
	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose
	that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$ and
	$\der{F}=\{\dder{F}_i\}_{i=0}^1$ are switches of $\dder{D}_0$ and
	$\dder{D}_1$. Then there is an abstraction equivalence between $\der{E}$ and $\der{F}$, whose first component is $\id{G_0}$.
\end{lemma}

\begin{proof}
	By definition of switch $\der{E}_0$ and $\der{F}_0$ have the same match $f_0\circ i_1$ and that $\der{E}_1$ and $\der{F}_1$ have the same comatch $g_1\circ i_0$. Thus we get the solid part of the diagram below
	\[\xymatrix@C=22pt{G_{0} && \ar@{>->}[ll]_{f_{\der{F},0}} D_{\der{F},0} \ar[rr]^{g_{\der{F},0}}&& G_{\der{F},1} && \ar@{>->}[ll]_{f_{\der{F},1}} D_{\der{F},1} \ar[rr]^{g_{\der{F},1}}&& G_{2}\\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}&& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar@{>->}[ll]_{l_1} \ar[r]^{r_1} & R_{1} \ar@/^.35cm/[drrr]|(.315)\hole_(.4){j_0} \ar[dr]|(.3)\hole_{h_{\der{E},0}} \ar@/_.35cm/[urrr]|(.315)\hole^(.4){a_0} \ar[ur]|(.3)\hole^{h_{\der{F},0}}& &L_{0} \ar@/_.35cm/[dlll]^(.4){j_1} \ar[dl]|(.3)\hole^{m_{\der{E},1}} \ar@/^.35cm/[ulll]_(.4){a_1} \ar[ul]|(.3)\hole_{m_{\der{F},1}}& K_{0} \ar[d]_{k_{\der{E},1}} \ar[u]^{k_{\der{F},1}}\ar@{>->}[l]_{l_{0}} \ar[rr]^{r_0} && R_{0} \ar[d]_{g_1\circ i_0} \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} && \ar[ll]^{f_{\der{E},0}} D_{\der{E},0} \ar@{.>}@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[rr]_{g_{\der{E},0}}&& G_{\der{E},1} \ar@{.>}[uu]^(.76){\psi_1}|(.45)\hole |(.55)\hole&& \ar@{>->}[ll]^{f_{\der{E},1}} D_{\der{E},1} \ar@{.>}@/_.5cm/[uu]_(.33){\phi_{1}}|\hole\ar[rr]_{g_{\der{E},1}}&& G_{2}\ar@{.>}@/_.5cm/[uu]_{\psi_2}}\]
